---
title: "DCM, ICM, and NM Modeling with EpiModel"
author: "David Wang"
date: "May 24, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(ape)
require(rCharts)
require(EpiModel)
require(reshape2)
require(knitr)
opts_chunk$set(comment = NA, results = "asis", comment = NA, tidy = F)
```

## Introduction

This brief analysis will conduct exercises on the three types of modeling techniques presented in the **EpiModel** package. They are the DCM (Deterministic Compartmental Model), ICM (Individual Compartmental Model or Agent-based Model), and Network Model. Within these models, there are several types of sub-models to choose from such as *SI*,*SIR*, and *SIS*. For each model, the analysis is broken down into three steps: (1) Input parameters, (2) Arrange output, and (3) Visualize.

## Simple DCM
1.) Input and initialize necessary parameters.
```{r dcm_input}
param <- param.dcm(inf.prob = 0.50, act.rate = 0.50, rec.rate=0.05)
init <- init.dcm(s.num = 999, r.num = 0, i.num = 1)
control <- control.dcm(type = "SIR", nsteps = 365)
mode <- dcm(param, init, control)
```

2.) Arrange output into dataframe and prepare it for custom visualization. 
```{r dcm_arrange}

results.df <- data.frame(mode)
prevNames <- c("s.num", "r.num", "i.num","num", "time")
incNames <- c("si.flow", "time")

prevDF <- results.df[prevNames]
incDF <- results.df[incNames]

# Convert prevDF cols into percentages of 'num'
percPrevDF <- prevDF
percPrevDF[c("s.num","r.num","i.num")] <- prevDF[c("s.num", "r.num", "i.num")] / prevDF$num
finalPrevNames <- c("s.num", "r.num", "i.num", "time")
finalPrevDF <- percPrevDF[finalPrevNames]

melt.prevDF <- melt(finalPrevDF, id="time")
melt.incDF <- melt(incDF, id="time")

# Clean dataframe
melt.prevDFsub <- subset(melt.prevDF, !is.na(value))
melt.prevDFsub <- subset(melt.prevDFsub, !is.infinite(value))

# Prevalence values are between 0 and 1 only
maxIsOne <- pmin(melt.prevDFsub$value, 1)
melt.prevDFsub$value <- maxIsOne
melt.prevDFsub$value[melt.prevDFsub$value <= 0] <- 0

```

3.) Create visualizations of Prevalence and Incidence versus Time.

Prevalence vs. Time
```{r dcm_vis_prev, echo=FALSE, cache=T, results='asis', comment=NA}
# Create linechart
p1 <- nPlot(value~time, group="variable", data = melt.prevDFsub, type = "lineChart")
p1$set(range = range(melt.prevDF$time))
p1$chart(yDomain = c(0, 1))
p1$xAxis(axisLabel = 'Timestep')
p1$yAxis(axisLabel = 'Prevalence')
p1$print('chart1', include_assets = T)

```
Incidence vs. Time
```{r dcm_vis_inc, echo=FALSE, cache=T, results='asis', comment=NA}
#melt.incDF$value <- as.integer(melt.incDF$value)
# melt.incDFsub <- subset(melt.incDF, !is.na(value))
p2 <- nPlot(value~time, group="variable", data = melt.incDF, type = "lineChart")
p2$set(range = range(melt.incDF$time))
p2$xAxis(axisLabel = 'Timestep')
p2$yAxis(axisLabel = 'Incidence')
#p2$chart(margin=list(left=300))
p2$print('chart2', include_assets = T)

```

