---
title: "Chapter3"
author: "David Wang"
date: "June 2, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(igraph)
```

### Graph Layouts
The graph layout is the placement of vertices and edges in space.
```{r}
library(sand)
g.l <- graph.lattice(c(5,5,5))

# Bring in example data and summarize
data(aidsblog)
aidsblog <- upgrade_graph(aidsblog)
summary(aidsblog)
```

Exploring a circular layout:
```{r}
igraph.options(vertex.size=3, vertex.label=NA,
               edge.arrow.size=0.5)

# Set the plot output arrangement (1 row with 2 elements)
par(mfrow=c(1,2))

plot(g.l, layout=layout.circle) #layout.circle sets the circular boundaries
title("5x5x5 Lattice")

plot(aidsblog, layout=layout.circle)
title("Blog Network")
```

Use `spring embedder` methods of graph drawing. (e.g. Fruchterman and Reingold method)
```{r}
# Plot same data again, but this time using a different layout method
par(mfrow=c(1,2))

plot(g.l, layout=layout.fruchterman.reingold)
title("5x5x5 Lattice")

plot(aidsblog, layout=layout.fruchterman.reingold)
title("Blog Network")
```

Use `energy placement` methods of graph drawing. (e.g. Kamada and Kawai)
```{r}
par(mfrow=c(1,2))

plot(g.l, layout=layout.kamada.kawai)
title("5x5x5 Lattice")

plot(aidsblog, layout=layout.kamada.kawai)
title("Blog Network")
```

Sometimes we need special formatting. Take this case with tree graphs for example:
```{r}
g.tree <- graph.formula(1-+2, 1-+3, 1-+4, 2-+5, 2-+6, 2-+7,
                        3-+8, 3-+9, 4-+10)

# Set the plot output arrangement (1 row with 3 elements)
par(mfrow=c(1,3))
igraph.options(vertex.size=30, edge.arrow.size=0.5,
               vertex.label=NULL)

# Plot g.tree three different ways
plot(g.tree, layout=layout.circle)
title("Cirlce Layout")

plot(g.tree, layout=layout.reingold.tilford(g.tree, circular=T))
title("Reingold & Tilford (Radial)")

plot(g.tree, layout=layout.reingold.tilford)
title("Reingold & Tilford (Layered)")
```

### Decorating Graph Layouts

One karate school splits into two karate schools with two leaders (Mr. Hi vs John A.)

Using karate example in network analysis example:

```{r}
library(igraphdata)
data(karate)

# We set the seed so that the layout is reprod.
set.seed(42)

l <- layout.kamada.kawai(karate)

# Plot without decoration (initially)
igraph.options(vertex.size=10)
par(mfrow=c(1,1))
plot(karate, layout=l, vertex.label=V(karate))

## Plot with decoration (the labels)
# Just isolate the leaders here
V(karate)$label <- sub("Actor ", "", V(karate)$name)

# Assign all vertices to be a circle
V(karate)$shape <- "circle"

# Now assign the leaders to be a different shape from a circle
V(karate)[c("Mr Hi", "John A")]$shape <- 'rectangle'

# Differentiate the two schools by color
V(karate)[Faction == 1]$color <- "red"
V(karate)[Faction == 2]$color <- "dodgerblue"

# Vertex area proportional to vertex strength
V(karate)$size <- 4*sqrt(graph.strength(karate))
V(karate)$size2 <- V(karate)$size * 0.5

# Weight edges by the number of common activities
E(karate)$width <- E(karate)$weight

# Color edges by within/between faction.
F1 <- V(karate)[Faction==1]
F2 <- V(karate)[Faction==2]

# School 1 to School 1 links are pink
E(karate)[F1 %--% F1]$color <- "pink"

# School 2 to School 2 links are lightblue
E(karate)[F2 %--% F2]$color <- "lightblue"

# School 1 to School 2 links are yellow
E(karate)[F1 %--% F2]$color <- "yellow"

## Offset vertex labels for smaller points so that it looks better
# If the size of the vertex is greater than 10, no need for offset, ow offset by 0.75
V(karate)$label.dist <-
        ifelse(V(karate)$size >= 10, 0, 0.75)

# Let's plot!
plot(karate, layout=l)
```

We can conduct a similar analysis using the `lazega_lawyer` dataset from the last chapter:
```{r}
library(sand)
data("lazega")
lazega <- upgrade_graph(lazega)

## Office location indicated by color
# Replaces all 3 different possible numbers in original vector with color strings
v.colors <- c("red", "dodgerblue", "goldenrod")[V(lazega)$Office]

# Type of practice indicated by vertex shape
v.shapes <- c("circle", "square")[V(lazega)$Practice]

# Vertex size proportional to years with firm
v.size <- 3.5*sqrt(V(lazega)$Years)

# Label vertices according to seniority
v.label <- V(lazega)$Seniority

set.seed(42)

l <- layout.fruchterman.reingold(lazega)
plot(lazega, layout = l, vertex.color = v.colors,
     vertex.shape = v.shapes, vertex.size = v.size,
     vertex.label = v.label)
```


## Visualizing Large Networks
As networks get larger, plots look more cluttered. E.g. blog example:
```{r}
library(sand)
fblog <- upgrade_graph(fblog)
summary(fblog)
party.names <- sort(unique(V(fblog)$PolParty))
# So much less clutter
party.names

# Use K-K method
set.seed(42)
l = layout.kamada.kawai(fblog)

# Convert graph object into factor (categories)
party.nums.f <- as.factor(V(fblog)$PolParty)

# Convert factors into numbers
party.nums <- as.numeric(party.nums.f)

par(mfrow=c(1,2))
plot(fblog, layout=l, vertex.label=NA,
     vertex.color=party.nums, vertex.size=3)
title("Kamada Kawai Method")

# DrL algorithm
set.seed(42)
l <- layout.drl(fblog)
plot(fblog, layout=l, vertex.size=5, vertex.label=NA,
     vertex.color=party.nums)
title("DrL Method")
```

Local vertex neighbors
```{r}
data(karate)
k.nbhds <- graph.neighborhood(karate, order=1)

# Find which ones have the largest clusters
sapply(k.nbhds, vcount)

# Plot the largest neighborhoods
k.1 <- k.nbhds[[1]]
k.34 <- k.nbhds[[34]]

par(mfrow=c(1,2))
plot(k.1, vertex.label=NA,
     vertex.color=c("red", rep("lightblue", 16)))
title("Mr. Hi")

plot(k.34, vertex.label=NA,
     vertex.color=c(rep("lightblue", 17), "red"))
title("John A")


```









